---
title: "IE 6200 Project - Section 9 - Group 1"
author: "Noah Klur, Jordan Lian"
date: "09-Dec-2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Collection

We chose to work with the crimedata library in R. To conduct our analysis, we used dplyr, ggplot2, lubridate, tidyr, and fitdistrplus.

```{r, , message=FALSE, warning=FALSE}
library(crimedata)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(fitdistrplus)
```

```{r, origin_crime, include=FALSE}
origin_crime <- get_crime_data()
```

``` {r ,, warning = FALSE, message = FALSE}
str(origin_crime)
origin_crime
```

## Data Visualization

We first looked at crimes per city before picking a few cities to look at. 

We created a table showing the total crimes per city in the year 2019. 

```{r crime_data, city_crime, message=FALSE, echo=FALSE, warning=FALSE}
crime_data <- origin_crime
city_count <- table(crime_data$city_name)
city_count
barplot(city_count, main="Total Number of Crimes per City in 2019",
        xlab="City",
        ylab="Frequency")
```

We then decided to look at crimes per month. We had to filter out the month only from each crime. We modified the date_single variable to help us do that. Then we created a separate data frame from these results.

```{r, crime_data, date_freq, message=FALSE, echo=FALSE, warning=FALSE}
crime_data$date_single <- format(as.POSIXct(crime_data$date_single, format='%Y/%m/%d %H:%M:%S'),format='%m')

date_freq <- as.data.frame(table(crime_data$date_single))
date_freq$avg <- as.numeric(as.character(date_freq$Freq)) / 15
colnames(date_freq) = c("Month", "Total_Freq", "Avg_Freq")
date_freq

barplot(date_freq$Avg_Freq, names.arg=date_freq$Month, main="Crime Frequency per Month",
        xlab="Month",
        ylab="Average Frequency per City")
```

From these results, we decided to look at New York, Chicago, Los Angeles, and Austin.


### New York
```{r, , message=FALSE, echo=FALSE, warning=FALSE}
# New York Crime Data set, filtered from crime_data
NY <- crime_data[crime_data$city_name == "New York",]
NY

  #group crimes in NY by month
NY_date_freq <- as.data.frame(table(NY$date_single))
colnames(NY_date_freq) = c("Month", "Month_Freq")
NY_date_freq

  #group total number of crimes in NY by offense code and create bar plot
NY_crime <- table(NY$offense_code)

barplot(NY_crime, main="New York Total Crime",
        xlab="Offense Code",
        ylab="Frequency")

  #create scatterplot of crimes in NY by month
legend_key <- theme(plot.title = element_text(family = "Helvetica", face = "bold", size = (15), hjust=0.5))
NY_scatter <- ggplot(NY_date_freq) +
  geom_point(mapping = aes(x = Month, y = Month_Freq))
print(NY_scatter + legend_key + labs(y="Crimes Each Month", title="New York Monthly Crimes"))
```

### Chicago
```{r, , message=FALSE, echo=FALSE, warning=FALSE}
# Chicago Crime Data set, filtered from crime_data
Chi <- crime_data[crime_data$city_name == "Chicago",]
Chi

  #group crimes in Chicago by month
Chi_date_freq <- as.data.frame(table(Chi$date_single))
colnames(Chi_date_freq) = c("Month", "Month_Freq")
Chi_date_freq

  #group total number of crimes in Chicago by offense code and create bar plot
Chi_crime <- table(Chi$offense_code)
barplot(Chi_crime, main="Chicago Total Crime",
        xlab="Offense Code",
        ylab="Frequency")

  #create scatterplot of crimes in Chicago by month
Chi_scatter <- ggplot(Chi_date_freq) +
  geom_point(mapping = aes(x = Month, y = Month_Freq))
print(Chi_scatter + legend_key + labs(y="Crimes each Month", title="Chicago Monthly Crimes"))
```

### Los Angeles
```{r, , message=FALSE, echo=FALSE, warning=FALSE}
# LA Crime Data set, filtered from crime_data
LA <- crime_data[crime_data$city_name == "Los Angeles",]
LA

  #group crimes in LA by month
LA_date_freq <- as.data.frame(table(LA$date_single))
colnames(LA_date_freq) = c("Month", "Month_Freq")
LA_date_freq

  #group total number of crimes in LA by offense code and create bar plot
LA_crime <- table(LA$offense_code)
barplot(LA_crime, main="Los Angeles Total Crime",
        xlab="Offense Code",
        ylab="Frequency")

  #create scatterplot of crimes in LA by month
LA_scatter <- ggplot(LA_date_freq) +
  geom_point(mapping = aes(x = Month, y = Month_Freq))
print(LA_scatter + legend_key + labs(y="Crimes each Month", title="Los Angeles Monthly Crime"))
```

### Austin
```{r, , message=FALSE, echo=FALSE, warning=FALSE}
# Austin Crime Data set, filtered from crime_data
Aus <- crime_data[crime_data$city_name == "Austin",]
Aus

  #group crimes in Austin by month
Aus_date_freq <- as.data.frame(table(Aus$date_single))
colnames(Aus_date_freq) = c("Month", "Month_Freq")
Aus_date_freq

  #group total number of crimes in Austin by offense code and create bar plot
Aus_crime <- table(Aus$offense_code)
barplot(Aus_crime, main="Austin Total Crime",
        xlab="Offense Code",
        ylab="Frequency")

  #create scatterplot of crimes in Austin by month
Aus_Scatter <- ggplot(Aus_date_freq) +
  geom_point(mapping = aes(x = Month, y = Month_Freq))
print(Aus_Scatter + legend_key + labs(y="Crimes Each Month", title="Austin Monthly Crimes"))

a <- data.frame(city_count)
colnames(a) = c("City", "Yearly_Crime")
a
Avg_Salary <- c(67462, 65883, 55198, 29481, 59255, 43573, 58385, 32131, 54700, 54505, 60762, 104552, 85562, 41107, 41625)
Population <- c(978908, 692600, 2693976, 670031, 909585, 152960, 3979576, 770517, 518012, 692587, 8336817, 881549, 753675, 300576, 548073)
data <- cbind(a,Avg_Salary)
Cor_data <- cbind(data,Population)
Cor_data
cor(Cor_data$`Yearly_Crime`, Cor_data$Avg_Salary)
cor_crime_salary <- ggplot(Cor_data) +
  geom_point(mapping = aes(x = Avg_Salary, y = Yearly_Crime))
print(cor_crime_salary + legend_key + labs(x="Average Yearly Salary", y="Yearly Crime", title="Correlation Plot: Avg. Yearly Salary vs Yearly Crime Rates"))
ggplot(data = Cor_data, aes(x=Cor_data$Avg_Salary, y=Cor_data$Yearly_Crime))+
  geom_point(aes(size=City, color=City))+
  labs(x="Average Salary", y="Yearly Crime")  
#correlation for yearly crime and population
cor(Cor_data$`Yearly_Crime`, Cor_data$Population)
cor_crime_population <- ggplot(Cor_data) +
  geom_point(mapping = aes(x = Population, y = Yearly_Crime))
print(cor_crime_population + legend_key + labs(x="Population", y="Yearly Crime", title="Correlation Plot: Avg. Population vs Yearly Crime Rates"))
ggplot(data = Cor_data, aes(x=Cor_data$Population, y=Cor_data$Yearly_Crime))+
  geom_point(aes(size=City, color=City))+
  labs(x="Population", y="Yearly Crime")
```

After getting visuals from all of the cities, we compared monthly crime from each of the 4 cities. We created a new data frame to plot it on the same graph.

### Grouped Bar Plot
```{r, city_crime, city_long, city_bar, messages=FALSE, echo=FALSE, warning=FALSE}
# Grouped Bar Plot
month_list <- c(1:12)
as.data.frame(month_list)
city_crime <- bind_cols(month_list, NY_date_freq$Month_Freq, Chi_date_freq$Month_Freq, LA_date_freq$Month_Freq, Aus_date_freq$Month_Freq)
colnames(city_crime) = c("Month", "NY_Month_Freq", "Chi_Month_Freq", "LA_Month_Freq", "Aus_Month_Freq")
city_crime <- as.data.frame(city_crime)
city_crime

# Modify Dataframe
city_long <- gather(city_crime, variable, value, -Month)
city_long

# Bar Plot
city_bar <- ggplot(city_long, aes(Month, value))
month_crimes <- city_bar + geom_bar(stat = "identity", aes(fill =  variable), position = "dodge") +
  xlab("Months") + ylab("Frequency") +
  ggtitle("Monthly Crimes by City")
print(month_crimes + legend_key + scale_fill_discrete(name = "City", labels = c("Austin", "Chicago", "Los Angeles", "New York")))
```

## Statistical Analysis
We conducted 3 tests.
 
1. Hypothesis Test: For our hypothesis test, we chose to do two tests of unequal variances - looking at New York and then Austin. For the first hypothesis test of unequal variances, we look at New York as the sample from the population being the United States.

- Step 1: Hypothesis:
- Null Hypothesis: $H_{0}$: $\sigma^2$ $\le$ $\sigma_{0}^2$
- Alternate Hypothesis: $H_{1}$: $\sigma^2$ > $\sigma_{0}^2$
- Step 2: Test Statistic
- $\chi^2_{calc} = \frac{(n-1)s^2}{\sigma_{0}^2}$
- For New York: n = 4720, $s^2 = 5436.424$, $\sigma_{0}^2 = 73.02411$
- For Austin: n = 878, $s^2 = 120.3333$, $\sigma_{0}^2 = 73.02411$
- Step 3: Rejection Region
- Step 4: Decision

2. Confidence Interval for New York
- $\alpha = 0.05$
- $\frac{(n-1)S^2}{\chi^2_{\alpha/2}} < \sigma^2 < \frac{(n-1)S^2}{\chi^2_{1-\alpha/2}}$

3. Probability Sampling
- $\alpha = 0.05$
- $Z = \frac{\overline{X} - \mu}{\frac{\sigma}{\sqrt{n}}}$

```{r, , echo=FALSE}

```

## Advanced Analytics
Below are some of the advanced analytics we chose to run for our project.

We started by doing descriptive statistics on the data, and the results are as follows:

```{r, , message=FALSE, echo=FALSE, warning=FALSE}
#Regression for advanced analytics
descdist(date_freq$Avg_Freq)
```

We then decided to fit the normal and lognormal distributions to the data and analyzing which seems to be better, as show 
below:

```{r, , message=FALSE, echo=FALSE, warning=FALSE}
fit_n <- fitdist(date_freq$Avg_Freq, "norm")
summary(fit_n)
fit_ln <- fitdist(date_freq$Avg_Freq, "lnorm")
summary(fit_ln)

gofstat((list(fit_n,fit_ln)))
```

We then ran statistical analysis such as a Q-Q plot, a P-P plot, empirical and theoretical CDFs, and histogram and theoretical densities, as shown below:

```{r, , message=FALSE, echo=FALSE, warning=FALSE}
par(mfrow=c(2,2))
plot.legend <- c("normal")
denscomp(list(fit_n), legendtext = plot.legend, xlab = "Average Crime", xlegend = 'topleft')
cdfcomp(list(fit_n), legendtext = plot.legend, xlab = "Average Crime",)
qqcomp(list(fit_n), legendtext = plot.legend, xlab = "Average Crime",)
ppcomp(list(fit_n), legendtext = plot.legend, xlab = "Average Crime",)
```

## Conclusion
We found this project very interesting and generated several conclusions. Firstly, we
found that New York, Chicago, and Los Angeles have significantly more yearly crimes than
the other 12 cities. This led us to analyze them more and try to find correlations in these
numbers. We thought of two possibilities and wanted to analyze yearly crime based on
average income and population size. After manipulating the data we found extremely weak
correlation between yearly crime and average yearly income. We then looked at yearly
crime and population size and found a very strong positive correlation. So, we can
confidently conclude that yearly crime rates are strongly correlated to population size for
cities.